---
title: "Homework 4"
author: "PSTAT 134"
format:
  html:
    toc: true
    toc-location: left
    toc-depth: 4
    embed-resources: true
    theme: simplex
    code-fold: show
editor: visual
execute:
  warning: false
  message: false
---

## Homework 4

**Note: If this is one of your two late homework submissions, please indicate below; also indicate whether it is your first or second late submission.**

\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

This homework assignment has you practice working with some text data, doing some natural language processing. I strongly advise using Lab 7 for assistance.

You also may need to use other functions. I encourage you to make use of our textbook(s) and use the Internet to help you solve these problems. You can also work together with your classmates. If you do work together, you should provide the names of those classmates below.

[Names of Collaborators (if any):]{.underline}

### Natural Language Processing

We'll work with the data in `data/spotify-review-data.csv`. This CSV file contains a total of 51,473 rows, each representing a unique user review for the Spotify application. The dataset has two columns:

- Review: This column contains the text of user reviews, reflecting their experiences, opinions, and feedback on the Spotify app.

- Sentiment label: This column categorizes each review as either "POSITIVE" or "NEGATIVE" based on its sentiment.

The data comes from this source at Kaggle: <https://www.kaggle.com/datasets/alexandrakim2201/spotify-dataset>

```{r}
library(tidyverse)
library(tidymodels)
library(reshape2)
library(wordcloud)
library(ggraph)
library(tidytext)
library(igraph)
library(ggplot2)
```


#### Exercise 1

Read the data into R (or Python, whichever you prefer).

Take a look at the distribution of `label`. Are there relatively even numbers of negative and positive reviews in the data set?

```{r}
spotify_data <- read_csv("data/spotify-review-data.csv")
table(spotify_data$label)
```
The numbers of positive and negative review are about even, there are approximately 25% more positive reviews than negative reviews.

#### Exercise 2

Take a random sample of $10,000$ reviews, stratified by `label`. All further exercises will be working with this smaller sample of reviews.

```{r}
set.seed(1)
spotify_split <- initial_split(spotify_data, prop = 10000 / nrow(spotify_data), strata = label)
spotify_sample <- training(spotify_split)
```


#### Exercise 3

Clean the reviews. Remove punctuation and convert the letters to lowercase.

Tokenize the reviews into words.

Remove stop words. (You can use any pre-made list of stop words of your choice.)

Verify that this process worked correctly.

```{r}
# I used AI here to help me with the punctuation
spotify_clean <- spotify_sample %>%
  mutate(id = row_number(),
         Review = str_replace_all(Review, '[[:punct:]]', ' '),
         Review = tolower(Review)) %>%
  unnest_tokens(word, Review) %>% 
  anti_join(stop_words) 

head(spotify_clean)
```


#### Exercise 4

Create a bar chart of the most commonly-occurring words (not including stop words).

Create bar charts of the most commonly-occurring words, broken down by `label`. What words are more common in positive reviews? What words are more common in negative reviews?

```{r}
# Common words
spotify_clean %>%
  count(word, sort = TRUE) %>%
  slice_max(n, n = 20) %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(n, word)) +
  geom_col() +
  labs(title = "Most Common Words Overall", y = NULL)

# Common words broken down by label
spotify_clean %>%
  count(label, word, sort = TRUE) %>%
  group_by(label) %>%
  slice_max(n, n = 15) %>%
  ungroup() %>%
  mutate(word = reorder_within(word, n, label)) %>%
  ggplot(aes(n, word, fill = label)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~label, scales = "free_y") +
  scale_y_reordered() +
  labs(title = "Most Common Words by Sentiment Label", y = NULL)
```
In negative reviews, "app" is much more common than in positive reviews. Other common negative words that are less present in positive review are "play", "ads", "premium", "playing", "update", "don", "fix", and "phone". Words that occur more in positive words are "love", "oy", "listening", "easy", "amazing", and "nice".

#### Exercise 5

Create a word cloud of the most commonly-occurring words overall, broken down by "positive" or "negative" sentiment (using the Bing sentiment lexicon).

```{r}
spotify_clean %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  acast(word ~ sentiment, value.var = "n", fill = 0) %>%
  comparison.cloud(colors = c("red", "blue"), max.words = 100)
```


#### Exercise 6

Calculate the tf-idf values for the words in the dataset.

Find the 30 words with the largest tf-idf values.

Find the 30 words with the smallest tf-idf values.


```{r}
spotify_tf_idf <- spotify_clean %>%
  count(id, label, word) %>%
  bind_tf_idf(term = word, document = id, n = n)

# 30 words with largest tf-idf
top_30_tfidf <- spotify_tf_idf %>%
  arrange(desc(tf_idf)) %>%
  head(30)
print(top_30_tfidf)

# 30 words with smallest tf-idf
bottom_30_tfidf <- spotify_tf_idf %>%
  arrange(tf_idf) %>%
  head(30)
print(bottom_30_tfidf)
```

#### Exercise 7

Find the 30 most commonly occurring bigrams.

Create graphs visualizing the networks of bigrams, broken down by `label`. That is, make one graph of the network of bigrams for the positive reviews, and one graph of the network for the negative reviews.

What patterns do you notice?

```{r fig.width=10, fig.height=8}
spotify_bigrams <- spotify_sample %>%
  unnest_tokens(bigram, Review, token = "ngrams", n = 2) %>%
  separate(bigram, c("word1", "word2"), sep = " ") %>%
  filter(!word1 %in% stop_words$word) %>%
  filter(!word2 %in% stop_words$word) %>%
  drop_na(word1, word2)

# 30 most commonly occurring bigrams
top_30_bigrams <- spotify_bigrams %>%
  unite(bigram, word1, word2, sep = " ") %>%
  count(bigram, sort = TRUE) %>%
  head(30)
print(top_30_bigrams, n=30)

a <- grid::arrow(type = "closed", length = unit(.15, "inches"))

# POSITIVE Bigram Graph
pos_graph <- spotify_bigrams %>% 
  filter(label == "POSITIVE") %>%
  count(word1, word2) %>% 
  filter(n > 5) %>% 
  graph_from_data_frame()

ggraph(pos_graph, layout = "fr") +
  geom_edge_link(aes(edge_alpha = n), show.legend = FALSE, arrow = a, end_cap = circle(.07, 'inches')) +
  geom_node_point(color = "purple", size = 4) +
  geom_node_text(aes(label = name), vjust = 1, hjust = 1, repel = TRUE) +
  theme_void() +
  labs(title = "Network of Bigrams: POSITIVE Reviews")

# NEGATIVE Bigram Graph
neg_graph <- spotify_bigrams %>% 
  filter(label == "NEGATIVE") %>%
  count(word1, word2) %>% 
  filter(n > 5) %>% 
  graph_from_data_frame()

ggraph(neg_graph, layout = "fr") +
  geom_edge_link(aes(edge_alpha = n), show.legend = FALSE, arrow = a, end_cap = circle(.07, 'inches')) +
  geom_node_point(color = "pink", size = 4) +
  geom_node_text(aes(label = name), vjust = 1, hjust = 1, repel = TRUE) +
  theme_void() +
  labs(title = "Network of Bigrams: NEGATIVE Reviews")
```

From the 30 most common bigrams, we see that a good amount are related to the features and business model of Spotify, such as "spotify premium", "free version", "sound quality", "play list", and "6 skips". There are also strong positive and negative bigrams in the top 30, with bigrams like "love spotify", "nice app", "amazing app" but also "1 star" and "stops playing".

When looking at the graph of positive bigrams, lots of the patterns revolve around usability and satisfaction. We see links like "variety", "range", and "wide", indicating that people enjoy Spotify's massive music library. "User" and "friendly", "quality" and "sound", "quality" and "experience" all support this pattern.

The graph of negative bigrams has lots of links about frictions while using the app. "Glitchy" and "frustrating", "phone" and "crashing", "takes" and "forever", "songs" and "disappear" are all telling of some issues within the service.

#### Exercise 8

Using the tokenized **words** and their corresponding tf-idf scores, fit a **linear support vector machine** to predict whether a given review is positive or negative.

- Split the data using stratified sampling, with 70% training and 30% testing;

- Drop any columns with zero variance;

- Fit a linear support vector machine using default values for any hyperparameters;

- Calculate the model **accuracy** on your testing data.

```{r}
# Pivoting tf-idf values into features
svm_data <- spotify_tf_idf %>%
  group_by(word) %>%
  filter(n() > 10) %>% # Drop heavily sparse words
  ungroup() %>%
  pivot_wider(id_cols = c(id, label), 
              names_from = word, 
              names_prefix = "word_",
              values_from = tf_idf, 
              values_fill = 0) %>%
  select(-id) %>%
  mutate(label = as.factor(label))

# Split data
set.seed(1)
svm_split <- initial_split(svm_data, prop = 0.70, strata = label)
svm_train <- training(svm_split)
svm_test <- testing(svm_split)

# Drop columns with zero variance
svm_recipe <- recipe(label ~ ., data = svm_train) %>%
  step_zv(all_predictors())

# Fit a linear SVM
svm_model <- svm_linear() %>%
  set_engine("kernlab") %>%
  set_mode("classification")

svm_wkflow <- workflow() %>%
  add_model(svm_model) %>%
  add_recipe(svm_recipe)

svm_fit <- svm_wkflow %>% fit(data = svm_train)

# Calculate model accuracy
svm_predictions <- predict(svm_fit, new_data = svm_test) %>%
  bind_cols(svm_test %>% select(label))

accuracy(svm_predictions, truth = label, estimate = .pred_class)
```
